PANCHANGE_data_preprocessing

### 2020-08-07 PANCHANGE_data_preprocessing2020-07-15.Rmd
used the .numeric tibble for creating new variables

# identify number of outliers
length(
  which(
  glad.dem.raw.id$Age_uncleaned > age_upper_limit |
    glad.dem.raw.id$Age_uncleaned < age_lower_limit
     )
)

# remove age outliers
glad.dem.raw.id <- glad.dem.raw.id %>%
  mutate(
    Age = 
      if_else(
        Age_uncleaned > age_upper_limit | 
        Age_uncleaned < age_lower_limit, 
        true = NA_real_,
        false = Age_uncleaned,
        missing = NA_real_
      )
  )

# Inspect colnames
colnames(glad.dem.raw.id)
```

@@

# Create categorical age groups per 10 years
glad.dem.raw.id <- glad.dem.raw.id %>%
  mutate(
    age_category_numeric =
      case_when(
          Age >= 16 & Age <= 18 ~ "1",
          Age >= 19 & Age <= 25 ~ "2",
          Age >= 26 & Age <= 35 ~ "3",
          Age >= 36 & Age <= 45 ~ "4",
          Age >= 46 & Age <= 55 ~ "5",
          Age >= 56 & Age <= 65 ~ "6",
          Age >= 66 & Age <= 70 ~ "7",
          Age >= 71 & Age <= 75 ~ "8",
          Age >= 76 & Age <= 80 ~ "9",
          Age >= 81 & Age <= 85 ~ "10",
          Age >= 86 & Age <= 90 ~ "11",
          Age >= 91 & Age <= 100 ~ "12",
          Age >= 101 & Age <= 120 ~ "13" # oldest person in the world is 117 years
                )
      )

glad.dem.raw.id <- glad.dem.raw.id %>%
    mutate(
      age_category =
        recode_factor(
          age_category_numeric, 
          "1" = "16 to 18 years",
          "2" = "19 to 25 years",
          "3" = "26 to 35 years",
          "4" = "36 to 45 years",
          "5" = "46 to 55 years",
          "6" = "56 to 65 years",
          "7" = "66 to 70 years",
          "8" = "71 to 75 years",
          "9" = "76 to 80 years",
          "10" = "81 to 85 years",
          "11" = "86 to 90 years",
          "12" = "91 to 100 years",
          "13" = "Above 100 years"
          )
    )

glad.dem.raw.id %>%
  freq(age_category)

@@

# Highest education
Proceed along qualifications fields. If field is not blank, take that as highest qual. If field is blank, check next field. If all fields are blank, NA
Assumes University degree > A level > NVQ > GCSE/O Level/CSE
```{r Highest education}
#create numeric version of the highest education variable
glad.dem.raw.id <- glad.dem.raw.id %>%
  mutate(
    highest_education_numeric =
      case_when(
        demographics.college_or_university_degree_numeric == "1" ~ 4,
        demographics.a_levelsas_levels_or_equivalent_numeric == "1" ~ 3,
        demographics.nvq_or_hnd_or_hnc_or_equivalent_numeric == "1" ~ 2,
        demographics.o_levelsgcses_or_equivalent_numeric == "1" ~ 1,
        demographics.cses_or_equivalent_numeric == "1" ~ 1)
    )

#recode the numeric version into a factor
glad.dem.raw.id <- glad.dem.raw.id %>%
  mutate(
    highest_education =
      recode_factor(
        highest_education_numeric,
        `1` = "GCSE/CSE",
        `2` = "NVQ",
        `3` = "A-levels",
        `4` = "University")
    )

glad.dem.raw.id %>%
  freq(highest_education)
```

Check highest education coding
```{r Highest education check}
glad.dem.raw.id %>%
  select(highest_education,
        demographics.college_or_university_degree_numeric,
        demographics.a_levelsas_levels_or_equivalent_numeric,
        demographics.nvq_or_hnd_or_hnc_or_equivalent_numeric,
        demographics.o_levelsgcses_or_equivalent_numeric,
        demographics.cses_or_equivalent_numeric
  )
```

```{r glad dem head}
head(glad.dem.raw.id)
```

```{r glad dem summary}
summarytools::dfSummary(
  glad.dem.raw.id,
  graph.col = F,
  valid.col = F,
  labels.col = F)
```

>>>>

# identify number of outliers
length(
  which(
  glad.dem.raw.id.numeric$Age_uncleaned > age_upper_limit |
    glad.dem.raw.id.numeric$Age_uncleaned < age_lower_limit
     )
)

# remove age outliers
glad.dem.raw.id.numeric <- glad.dem.raw.id.numeric %>%
  mutate(
    Age = 
      if_else(
        Age_uncleaned > age_upper_limit | 
        Age_uncleaned < age_lower_limit, 
        true = NA_real_,
        false = Age_uncleaned,
        missing = NA_real_
      )
  )

# Inspect colnames
colnames(glad.dem.raw.id.numeric)
```

@@

# Create categorical age groups per 10 years
glad.dem.raw.id.numeric <- glad.dem.raw.id.numeric %>%
  mutate(
    age_category_numeric =
      case_when(
          Age >= 16 & Age <= 18 ~ "1",
          Age >= 19 & Age <= 25 ~ "2",
          Age >= 26 & Age <= 35 ~ "3",
          Age >= 36 & Age <= 45 ~ "4",
          Age >= 46 & Age <= 55 ~ "5",
          Age >= 56 & Age <= 65 ~ "6",
          Age >= 66 & Age <= 70 ~ "7",
          Age >= 71 & Age <= 75 ~ "8",
          Age >= 76 & Age <= 80 ~ "9",
          Age >= 81 & Age <= 85 ~ "10",
          Age >= 86 & Age <= 90 ~ "11",
          Age >= 91 & Age <= 100 ~ "12",
          Age >= 101 & Age <= 120 ~ "13" # oldest person in the world is 117 years
                )
      )

glad.dem.raw.id.numeric <- glad.dem.raw.id.numeric %>%
    mutate(
      age_category =
        recode_factor(
          age_category_numeric, 
          "1" = "16 to 18 years",
          "2" = "19 to 25 years",
          "3" = "26 to 35 years",
          "4" = "36 to 45 years",
          "5" = "46 to 55 years",
          "6" = "56 to 65 years",
          "7" = "66 to 70 years",
          "8" = "71 to 75 years",
          "9" = "76 to 80 years",
          "10" = "81 to 85 years",
          "11" = "86 to 90 years",
          "12" = "91 to 100 years",
          "13" = "Above 100 years"
          )
    )

glad.dem.raw.id.numeric %>%
  freq(age_category)

@@

# Highest education
Proceed along qualifications fields. If field is not blank, take that as highest qual. If field is blank, check next field. If all fields are blank, NA
Assumes University degree > A level > NVQ > GCSE/O Level/CSE
```{r Highest education}
#create numeric version of the highest education variable
glad.dem.raw.id.numeric <- glad.dem.raw.id.numeric %>%
  mutate(
    highest_education_numeric =
      case_when(
        demographics.college_or_university_degree_numeric == "1" ~ 4,
        demographics.a_levelsas_levels_or_equivalent_numeric == "1" ~ 3,
        demographics.nvq_or_hnd_or_hnc_or_equivalent_numeric == "1" ~ 2,
        demographics.o_levelsgcses_or_equivalent_numeric == "1" ~ 1,
        demographics.cses_or_equivalent_numeric == "1" ~ 1)
    )

#recode the numeric version into a factor
glad.dem.raw.id.numeric <- glad.dem.raw.id.numeric %>%
  mutate(
    highest_education =
      recode_factor(
        highest_education_numeric,
        `1` = "GCSE/CSE",
        `2` = "NVQ",
        `3` = "A-levels",
        `4` = "University")
    )

glad.dem.raw.id.numeric %>%
  freq(highest_education)
```

Check highest education coding
```{r Highest education check}
glad.dem.raw.id.numeric %>%
  select(highest_education,
        demographics.college_or_university_degree_numeric,
        demographics.a_levelsas_levels_or_equivalent_numeric,
        demographics.nvq_or_hnd_or_hnc_or_equivalent_numeric,
        demographics.o_levelsgcses_or_equivalent_numeric,
        demographics.cses_or_equivalent_numeric
  )
```

```{r glad dem head}
head(glad.dem.raw.id.numeric)
```

```{r glad dem summary}
summarytools::dfSummary(
  glad.dem.raw.id.numeric,
  graph.col = F,
  valid.col = F,
  labels.col = F)
```
